### 2022-11-08 paper/header.tex
Add files via upload
copied from sharelatex

% **************GENERATED FILE, DO NOT EDIT**************

\title{Julia for HPC: In Situ data Analysis with Julia for Climate Simulations at Large Scale}

\author[1]{Li Tang}
\author[2]{Soumya Dutta}
\author[1]{Natalie Klein}
\author[3]{Wayne Yu Wang}
\author[1]{Jonathan David Wolfe}
\author[1]{Luke Van Roekel}
\author[4]{Nathan Urban}
\author[1]{Ayan Biswas}
\author[1]{Earl Lawrence}
\affil[1]{Computer, Computational, and Statistical Sciences (CCS), Los Alamos National Laboratory}
\affil[2]{Department of Computer Science and Engineering, Indian Institute of Technology Kanpur}
\affil[3]{Department of Statistics, University of Michigan}
\affil[4]{Computational Science Initiative, Brookhaven National Laboratory}

\keywords{Julia, High-performance computing, In Situ data analysis, Scientific computing, Climate simulations, E3SM}

\hypersetup{
pdftitle = {In-Situ E3SM},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {Julia, In-Situ data analysis, Scientific computing, Climate simulations},
}